\documentclass[12pt]{article}
\usepackage[margin=0.75in]{geometry}
\usepackage{indentfirst}
\usepackage{listings}
\usepackage{amssymb}

\title{\bf Algorithms Homework 2}
\author{Kaitlin Poskaitis, Joshua Matthews, Russ Frank, and Matt Robinson}
\date{}

\begin{document}

\maketitle

\begin{enumerate}
%Question 1
\item Suppose that you are given an $k-sorted$ array, in which no element is
    farther than $k$ positions away from its final (sorted) position. Give an
    algorithm which will sort such an array. Prove its correctness. Analyse its
    running time. Note: your algorithm should run faster than $\Theta(nlogn)$,
    that is it should take advantage of the fact that the array is almost
    sorted.


\textbf{Algorithm:}

\begin{enumerate}
    \item Given a $k-sorted$ list of $n$ elements, split the list into
        $\frac{n}{k}$ lists of $k$ elements.

    \item Perform a merge sort on each of the sub-lists. \hfill $O(\frac{n}{k} *
        klog(k)) = O(nlogk)$

    \item At this point each element will still be no greater than $k$ away from
        its sorted position.

    \item As in mergesort, merge all of these sorted sub-lists back into a
        single list. Each of these merges can move each element at most twice as
        far from it's initial position, but each successive merge will cover
        twice as many elements and eventually place each element in it's correct
        position. \hfill $O(n)$

    \item Final efficiency is $O(nlogk) + O(n) = O(nlogk)$


\end{enumerate}


%Question 2
\item Consider once again the $k-sorted$ array of the previous problem. Show
    that any comparison based algorithm for sorting an almost sorted array makes
    $\Omega(nlogk)$ comparisons.

%Question 3
\item Prove that if an $n$ node graph has two of the following properties, it has the
third:
\begin{itemize}
    \item It has $n-1$ edges.
    \item It is connected
    \item It is acyclic
\end{itemize}

In order to prove this, we will use 2 lemmas whose proofs are trivial:\\
1. The fewest number of edges needed to connect $n$ nodes is $n-1$.\\
2. The fewest number of edges between $n$ nodes that must result in a cycle is
$n$.

Proof 1: If an $n$ node graph has $n-1$ edges and is connected, then it is
acyclic.\\
Because $n-1$ edges is the minimum number of edges to connect $n$ nodes, then
the graph must be acyclic because if it were cyclic, then it would not contain
the minimum number of edges to connect the nodes.

Proof 2: If an $n$ node graph is connected and acyclic, then it has $n-1$
edges.\\
If a graph is connected and acyclic, this means that it must have at least the
minimum number of edges to connect the nodes, but less than the number of edges
where a cycle is require. Therefore, it must have at least $n-1$ edges and less
than $n$ edges by the lemmas. Therefore, it must have $n-1$ edges.

Proof 3: If an $n$ node graph has $n-1$ edges and is acyclic, then it is
connected.\\
Let the first $n-2$ edges be used to connect $n-1$ nodes, which will not create
a cycle. The last edge can be used to connect the final node, or must create a
cycle with one of the other nodes. Therefore, it must connect the final node,
creating a connected graph.

%Question 4
\item Define a subtree of be any connected subgraph of a tree.
    \begin{itemize}
        \item Prove that the number of subtrees of a complete binary tree is not
            polynomial in the number of nodes.
        \item Give an example of a class of trees ${T_n}$ where the number of
            subtrees is a polynomial in the number of nodes.
    \end{itemize}

%Question 5
\item Let $F_i$ be the $i$th Fibonacci number. Show that $F_{n+2} = 1 +
    \sum_{i=0}^n F_i$.

%Question 6
\item Show that if you have a polynomial time algorithm for Hamiltonian Path,
    that you have a polynomial time algorithm for sorting.

    Proof:\\
    Assume there is a polynomial time algorithm for Hamiltonian Path.\\
    Merge sort is a polynomial sorting algorithm.\\
    QED

%Question 7
\item The Bounded Degree Spanning Tree (BDST) problem is the following:\\
    Input: Graph $G$ and integer $k$.\\
    Output: Yes, if $G$ has a spanning tree where every node has a degree of at
    most $k$, No, otherwise.\\
    Suppose there is no polynomial time algorithm for Hamilonian Path. Show that
    there is no polynomial time algorithm for BDST.

\end{enumerate}

\end{document}
